---
title: "W1. Phenotype data"
author: "Bora Kim"
date: "March 4, 2021"
output: html_document
---

This is meant for sharing dataset and script of the publication "Strigolactone_structural specificity in microbiome recruitment in rice, 2021". This markdown contains the process of phenotypic data: Biomass and levels of strigolactones (SLs) from 16 genotypes of rice grown two natural soils for 31 days. 




## 1. Getting strarted

Load required dataset (samples names on row, phenotypic observation on colunm). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
load("W1_phenotype_data_image.Rdata")
meta[1:5,1:5]
```



Load required packages for analysis

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(FSA) 
library(rcompanion) 
library(dplyr)
library(ggfortify)
library(tibble)
library(reshape2)
```



## 2. Comparing biomass (Fig S1)


To handle data easier, grouping information and biomass were subsetted from initial meta data
```{r echo=TRUE}

bm<-subset(meta[,c(8,1:4)],Compartment=="Root")
bm$Genotype2<-factor(bm$Genotype, 
                     c("IAC165", "IAC1246", "GWD", "Dullo","Bina","Sonk","Kinko", 
                       "TN1","Bhas","SC","Shiokari","d14","d3","d10","d17","d27")) #change order of genotype for presentation on plot later

```


### 2.1. Kruskal-Wallis test
Test the biomass by soil type and genotype in each soil using non-parametric Kruskal-Wallis test together with P value correction using 
Benjamini-Hochberg method. 


```{r echo=TRUE, message=FALSE, warning=FALSE}
bm_kw<-kruskal.test(bm$total_fresh_biomass_g_per_plant~bm$Soil, data=bm)
bm_kw
```

Create a result table combining all results

```{r echo=TRUE, message=FALSE, warning=FALSE}
bm_kw_res<-as.data.frame(matrix(NA, 1, 4))

bm_kw_res[,1]<-bm_kw$statistic
bm_kw_res[,2]<-bm_kw$parameter
bm_kw_res[,3]<-bm_kw$p.value
bm_kw_res[,4]<-p.adjust(bm_kw$p.value,method="BH" )
  
colnames(bm_kw_res)<-c("chi_squared", "Df", "P value", "adjusted P value")
rownames(bm_kw_res)<-"Total biomass_by_soil" 

bm_kw_res
```

### 2.3. Duun test
Compare biomass among genotypes in each soil, therefore subset the biomass from each soil

```{r echo=TRUE, message=FALSE, warning=FALSE}
fobm_dunn<-subset(bm, Soil=="Forest")
PT<-dunnTest(fobm_dunn$total_fresh_biomass_g_per_plant~Genotype, data=fobm_dunn, method = "bh")
fobm_dunn_res<-PT$res
fobm_dunn_res_letter<-cldList(comparison = fobm_dunn_res$Comparison,p.value = fobm_dunn_res$P.adj,threshold  = 0.05)
fobm_dunn_res[1:5,1:4]
fobm_dunn_res_letter
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
fibm_dunn<-subset(bm, Soil=="Forest")
PT<-dunnTest(fibm_dunn$total_fresh_biomass_g_per_plant~Genotype, data=fibm_dunn, method = "bh")
fibm_dunn_res<-PT$res
fibm_dunn_res_letter<-cldList(comparison = fibm_dunn_res$Comparison,p.value = fibm_dunn_res$P.adj,threshold  = 0.05)
fibm_dunn_res[1:5,1:4]
fibm_dunn_res_letter
```


### 2.4. Plot biomass


```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(bm) + 
  geom_boxplot(aes(x=Genotype2, y=total_fresh_biomass_g_per_plant, fill=Soil), outlier.colour = NA)+ 
  labs(x="", y = "Total fresh biomass g/plant") + scale_fill_manual(values=c("#a6611a","#018571"))+
  geom_point(aes(x=Genotype2, y=total_fresh_biomass_g_per_plant, fill=Soil), alpha = 0.3, shape = 21, 
             position = position_jitterdodge())+
  theme(axis.text.x = element_text(size = 13, colour = "black",angle=90, hjust=1), 
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(face = "bold", size = 13, vjust = 3), 
        legend.text = element_text(size = 13), legend.title = element_text(size = 13,face = "bold"))
```




## 3. Comparing strigolactone level (figure 1)

As SLs were not detected in field soil, further SL-related studies are only samples obtained from forest soil

```{r echo=TRUE}
foSLs<-subset(meta, Soil_compartment=="Fo_RS"&SL_analysis=="yes") 
```


### 3.1. Principle component analysis on three endogenous SLs

```{r echo=TRUE, message=FALSE, warning=FALSE}
foSLs<-subset(meta, Soil_compartment=="Fo_RS"&SL_analysis=="yes") 
foSLs<-subset(meta, Soil_compartment=="Fo_RS"&SL_analysis=="yes") 
foSLs_pca<-prcomp(foSLs[11:13], scale. = T)


autoplot(foSLs_pca, data = foSLs_pca, loadings = TRUE, 
         loadings.label = TRUE, loadings.label.size = 4)



```


### 3.2. Dunn test 

First, remove some genotypes due to zeros in all replicates (non-detectable) which could make problem during Dunn test.

```{r echo=TRUE}
foSLs_dunn<-foSLs %>% filter(!Genotype%in%c('Kinko','TN1','Bhas','SC','d10','d17','d27')) 
foSLs_dunn<-foSLs_dunn[,c(11:13,1)]
foSLs_dunn[1:5,1:4]

```

Perfom Dunn test using loop. As level of MeO5DS was not significantly different among genotype, it was excluded. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

indices=2 
Z<-as.data.frame(matrix(NA, 36, indices)) #results list =36
P.unadj<-as.data.frame(matrix(NA, 36, indices)) #results list =36
P.adj<-as.data.frame(matrix(NA, 36, indices)) #results list =36
foSLs_dunn_letter<-as.data.frame(matrix(NA, 9, indices)) #results list =9

for(i in 1:indices) {
  PT<-dunnTest(foSLs_dunn[,i]~Genotype, data=foSLs_dunn, method = "bh")
  Z[,i]<-PT$res$Z
  P.unadj[,i]<-PT$res$P.unadj
  P.adj[,i]<-PT$res$P.adj
  PT2<-PT$res
  cl<-cldList(comparison = PT2$Comparison, p.value = PT2$P.adj,threshold  = 0.05)
  foSLs_dunn_letter[,i]<-cl$Letter
}

foSLs_dunn_res<-cbind(Z,P.unadj, P.adj)
rownames(foSLs_dunn_res)<- PT$res$Comparison
colnames(foSLs_dunn_res)<-c("Z_orobanchol","Z_4DO",
                             "unadjusted.P_orobanchol","unadjusted.P_4DO",
                             "adjusted.P_orobanchol","adjusted.P_4DO")
rownames(foSLs_dunn_letter) <- cl$Group
colnames(foSLs_dunn_letter) <- colnames(foSLs_dunn[1:indices])

foSLs_dunn_res[1:5, 1:6]
foSLs_dunn_letter

```



### 3.3. Plot SLs
```{r echo=TRUE}
foSLs$Genotype<-droplevels(as.factor(foSLs$Genotype))
foSLs$Genotype2<-factor(foSLs$Genotype, 
                         c("IAC165", "IAC1246", "GWD", "Dullo","Bina","Sonk",
                           "Kinko", "TN1","Bhas","SC","Shiokari","d3","d14","d10","d17","d27"))

ggplot(foSLs)+
  geom_boxplot(aes(x=Genotype2, y=X4DO_pmol_g), outlier.colour = NA)+labs(x="Rice genotype", y = "4DO pmol/gFW") + 
  geom_point(aes(x=Genotype2, y=X4DO_pmol_g), alpha = 0.3, shape = 21) +
  theme(axis.text.x = element_text(size = 13, colour = "black",angle=90, hjust=1), 
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(face = "bold", size = 13, vjust = 3), 
        legend.text = element_text(size = 13), legend.title = element_text(size = 13,face = "bold"))



```





## version

```{r echo=TRUE}
sessionInfo()
```


